---
title: "Group_Project"
author: "Feiya Wang"
date: "11/11/2019"
output: html_document
---
***
## Group Project--Species Distribution Model

***
It is a Group projetc worked by Majo, Frank and Feiya




>Introduction: short paragraph and an image

>Packages to be used: install.packages(c(’raster’, ’rgdal’, ’dismo’, ’rJava’)) 
Raster: Geographic Data Analysis and Modeling
Rgdal: Bindings for the 'Geospatial' Data Abstraction Library
Dismo: Species Distribution Modeling
rJava: Low-Level R to Java Interface

# Data preparation - due Monday 18th

>Species occurrence data

```{r installing packages}
install.packages("raster")
install.packages("dismo")
install.packages("rgdal")
install.packages("rJava")
```

Importing data from a database

```{r loading package dismo}
library(dismo)
```


```{r}
```


```{r}
acaule <- gbif("solanum", "acaule*", geo=FALSE)
```

load the saved S. acaule data
```{r}
data("acaule")
```
how many rows and colums?

```{r}
dim(acaule)
```

Select the records that have longitude and latitude data


```{r}
colnames(acaule)
```


Subetting the data that have longitude and latitude


```{r}
acgeo <- subset(acaule, !is.na(lon) & !is.na(lat))
dim(acgeo)
```
show some values

```{r}
acgeo[1:4, c(1:5,7:10)] #removing locality
```

>Importing data: Species occurrence data, plot map- majo

```{r}
library(maptools)
```


```{r}
data(wrld_simpl)
plot(wrld_simpl, xlim=c(-80,70), ylim=c(-80,30), axes=TRUE, col="light yellow")

# restore the box around the map
box()

# add the points
points(acgeo$lon, acgeo$lat, col='orange', pch=20, cex=0.75)
# plot points again to add a border, for better visibility
points(acgeo$lon, acgeo$lat, col='red', cex=0.75)
```

```{r}
 r <- raster(system.file("external/test.grd", package="raster"))

```


```{r}
getData('GADM', country='FRA', level=1)

```

>Data cleaning-majo

```{r}
acaule[c(303,885),1:10]
```


```{r}
lonzero = subset(acgeo, lon==0)
lonzero[, 1:13]
```


>Duplicate records - Frank

>cross-checking


>Absence and background points: Some of the early species distribution model algorithms, such as Bioclim and Domain only use ’presence’ data in the modeling process. Other methods also use ’absence’ data or ’background’ data--Feiya 
 
# Environmental data Due 25th

>Raster data

>Extracting values from rasters

# Model fitting, prediction, and evaluation
>Model fitting

>Model prediction

>Model evaluation

>Types of models

# Generalized Linear Models: example









```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

