---
title: "Group_Project"
author: "Feiya Wang"
date: "11/11/2019"
output: html_document
---
***
## Group Project--Species Distribution Model

***
It is a Group projetc worked by Majo, Frank and Feiya




>Introduction: short paragraph and an image

>Packages to be used: install.packages(c(’raster’, ’rgdal’, ’dismo’, ’rJava’)) 
Raster: Geographic Data Analysis and Modeling
Rgdal: Bindings for the 'Geospatial' Data Abstraction Library
Dismo: Species Distribution Modeling
rJava: Low-Level R to Java Interface

# Data preparation - due Monday 18th



>Species occurrence data

>Importing data: Species occurrence data, plot map- majo

>Data cleaning-majo

>Duplicate records - Frank

>cross-checking


>Absence and background points: 

Some of the early species distribution model algorithms, such as Bioclim and Domain only use ’presence’ data in the modeling process. Other methods also use ’absence’ data or ’background’ data.

First, there is a common methods to delete all NA values from datae frame. We can use na.omit()function to delete the rows that contains the NA data.
```{r}

na.omit(df)
```

The dismo package also have some function to delete NA data from the data frame.
For example, raster()
 
# Environmental data Due 25th

>Raster data

>Extracting values from rasters

# Model fitting, prediction, and evaluation
>Model fitting

>Model prediction

>Model evaluation

>Types of models

# Generalized Linear Models: example









```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

